Results Text

### 2024-03-19 Fish Data Analysis/Paper_Figures.qmd
Formatting fixes and table

```{r}
fish_comp_data <- read.csv("Data/Fish_Comp_Values_3D.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                           ### Used to remove every third point to reduce autocorrelation
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"),
                                  Date = paste(Year,Month,Day,Trial, sep = "_"),
                                  Fish_ID = paste(Year,Month,Day,Trial,Fish1, sep = "_"),
                                  Comp_ID = paste(Year,Month,Day,Trial,Fish1,Fish2, sep = "_")) %>%
                           group_by(Comp_ID) %>%
                           #filter(row_number() %% 3 == 1) %>%
                           ungroup()

@@

#close
fish_comp_data_ring_angles <- fish_comp_data %>% filter(Distance >= min_dist & Distance <= max_dist) %>%
                                                 mutate(Dist_Cat = ifelse(Distance <= middle_dist, "Close", "Far")) %>%
                                                 # group_by(Date,Year,Month,Day,Trial,
                                                 #          Flow,Ablation,Darkness,Dist_Cat,Flow_Ablation_Darkness,Fish1,Tailbeat_Num) %>%
                                                 # summarise(Mean_Angle = rad_ang_mean(Angle),
                                                 #           #Should be the same number each time
                                                 #           Fish1_Yaw_Heading_PS = mean(Fish1_Yaw_Heading_PS, rm.na = T)) %>%
                                                 # ungroup()
                                                   mutate(Mean_Angle = Angle)

@@

nn_heading_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Mean_Angle,pi/6)) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness,Dist_Cat) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() 
```

@@

perp_angle_df <- fish_data_nn_normal %>% filter(Mean_Angle >= deg2rad(75) & Mean_Angle <= deg2rad(105))

perp_angle_df_neg <- fish_data_nn_normal %>% filter(Mean_Angle >= deg2rad(-105) & Mean_Angle <= deg2rad(-75)) %>% 
                                                   mutate(Fish1_Yaw_Heading_PS = -1*Fish1_Yaw_Heading_PS)

@@

heading_PS_graph_close <- ggplot(perp_angle_df %>% filter(Dist_Cat == "Close"),
                                 aes(x = interaction(Ablation, Darkness, sep =", "), y = rad2deg(Fish1_Yaw_Heading_PS),
                          fill = interaction(Ablation, Darkness, sep =", "), color = interaction(Ablation, Darkness, sep =", ")))+
  stat_dots(side = "both", binwidth = unit(dot_binwidth*0.375, "npc")) +
  geom_point(aes(y = rad2deg(mean_turn)), size = 5, color = "black") +
  geom_errorbar(aes(ymin=rad2deg(mean_turn-sd_turn), ymax=rad2deg(mean_turn+sd_turn)), width=.2, color = "black")+
  theme_light()+
  scale_fill_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_color_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  theme(legend.position = "none")+
  xlab("")+
  ylab("Turning Angle (Degrees / s)")+
  scale_y_continuous(breaks = seq(-180,180, 60))+
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", hide.ns = TRUE) +
  theme(axis.text.x = element_text(size = text_size),
        axis.text.y = element_text(size = text_size),
        axis.title.y = element_text(size = text_size))+
  scale_x_discrete(labels=function(x){sub(",\\s", "\n", x)})

@@

heading_PS_graph_far <- ggplot(perp_angle_df %>% filter(Dist_Cat == "Far"),
                                 aes(x = interaction(Ablation, Darkness, sep =", "), y = rad2deg(Fish1_Yaw_Heading_PS),
                          fill = interaction(Ablation, Darkness, sep =", "), color = interaction(Ablation, Darkness, sep =", ")))+
  stat_dots(side = "both", binwidth = unit(dot_binwidth*0.375, "npc")) +
  geom_point(aes(y = rad2deg(mean_turn)), size = 5, color = "black") +
  geom_errorbar(aes(ymin=rad2deg(mean_turn-sd_turn), ymax=rad2deg(mean_turn+sd_turn)), width=.2, color = "black")+
  theme_light()+
  scale_fill_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_color_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  theme(legend.position = "none")+
  xlab("")+
  ylab("Turning Angle (Degrees / s)")+
  scale_y_continuous(breaks = seq(-180,180, 60))+
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", hide.ns = TRUE) +
  theme(axis.text.x = element_text(size = text_size),
        axis.text.y = element_text(size = text_size),
        axis.title.y = element_text(size = text_size))+
  scale_x_discrete(labels=function(x){sub(",\\s", "\n", x)})

@@

heading_PS_mega_plot <- annotate_figure(heading_PS_mega_plot, 
                top = text_grob("Near (< 2.5 BL)                                          Far (> 2.5 BL)",
                                size = 24))

@@

single_plot <- ggplot(single_perp_angle_graphing, aes(x = Fish1_Yaw_Heading_PS, fill = Label))+
  geom_histogram(aes(y=0.5*..density..),binwidth=1,position='identity', alpha = 0.5)+
  theme_classic()+
  scale_fill_manual(values = c("#CE2414","#138EC2"))+
  ylab("Proportion of Values")+
  xlab("Turning Angle (Degrees / s)")+
  xlim(-5,5)

>>>>

```{r}
fish_comp_data <- read.csv("Data/Fish_Comp_Values_3D.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"),
                                  Date = paste(Year,Month,Day,Trial, sep = "_"),
                                  Fish_ID = paste(Year,Month,Day,Trial,Fish1, sep = "_"),
                                  Comp_ID = paste(Year,Month,Day,Trial,Fish1,Fish2, sep = "_"))

@@

#close
fish_comp_data_ring_angles <- fish_comp_data %>% filter(Distance >= min_dist & Distance <= max_dist) %>%
                                                 mutate(Dist_Cat = ifelse(Distance <= middle_dist, "Close", "Far"))

@@

nn_heading_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Angle,pi/6)) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness,Dist_Cat) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() 
```

@@

perp_angle_df <- fish_data_nn_normal %>% filter(Angle >= deg2rad(75) & Angle <= deg2rad(105))

perp_angle_df_neg <- fish_data_nn_normal %>% filter(Angle >= deg2rad(-105) & Angle <= deg2rad(-75)) %>% 
                                                   mutate(Fish1_Yaw_Heading_PS = -1*Fish1_Yaw_Heading_PS)

@@

heading_PS_graph_close <- ggplot(perp_angle_df %>% filter(Dist_Cat == "Close"),
                                 aes(x = interaction(Ablation, Darkness, sep =", "), y = rad2deg(Fish1_Yaw_Heading_PS),
                          fill = interaction(Ablation, Darkness, sep =", "), color = interaction(Ablation, Darkness, sep =", ")))+
  stat_dots(side = "both", binwidth = unit(dot_binwidth*0.375, "npc")) +
  geom_point(aes(y = rad2deg(mean_turn)), size = 5, color = "black") +
  geom_line(aes(y = rad2deg(mean_turn), group = 1), size = 1, color = "black") +
  geom_errorbar(aes(ymin=rad2deg(mean_turn-sd_turn), ymax=rad2deg(mean_turn+sd_turn)), width=.2, color = "black")+
  theme_light()+
  scale_fill_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_color_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  theme(legend.position = "none")+
  xlab("")+
  ylab("Focal Fish Turning Angular Speed (Degrees / s)")+
  scale_y_continuous(breaks = seq(-180,180, 60))+
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", hide.ns = TRUE) +
  theme(axis.text.x = element_text(size = text_size),
        axis.text.y = element_text(size = text_size),
        axis.title.y = element_text(size = text_size))+
  scale_x_discrete(labels=function(x){sub(",\\s", "\n", x)})

@@

heading_PS_graph_far <- ggplot(perp_angle_df %>% filter(Dist_Cat == "Far"),
                                 aes(x = interaction(Ablation, Darkness, sep =", "), y = rad2deg(Fish1_Yaw_Heading_PS),
                          fill = interaction(Ablation, Darkness, sep =", "), color = interaction(Ablation, Darkness, sep =", ")))+
  stat_dots(side = "both", binwidth = unit(dot_binwidth*0.375, "npc")) +
  geom_point(aes(y = rad2deg(mean_turn)), size = 5, color = "black") +
   geom_line(aes(y = rad2deg(mean_turn), group = 1), size = 1, color = "black") +
  geom_errorbar(aes(ymin=rad2deg(mean_turn-sd_turn), ymax=rad2deg(mean_turn+sd_turn)), width=.2, color = "black")+
  theme_light()+
  scale_fill_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_color_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  theme(legend.position = "none")+
  xlab("")+
  ylab("Focal Fish Turning Angular Speed (Degrees / s)")+
  scale_y_continuous(breaks = seq(-180,180, 60))+
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", hide.ns = TRUE) +
  theme(axis.text.x = element_text(size = text_size),
        axis.text.y = element_text(size = text_size),
        axis.title.y = element_text(size = text_size))+
  scale_x_discrete(labels=function(x){sub(",\\s", "\n", x)})

@@

heading_PS_mega_plot <- annotate_figure(heading_PS_mega_plot, 
                top = text_grob("  Near (< 2.5 BL)                                                Far (> 2.5 BL)",
                                size = 24))

@@

single_plot <- ggplot(single_perp_angle_graphing, aes(x = rad2deg(Fish1_Yaw_Heading_PS), fill = Label))+
  geom_histogram(aes(y=0.5*..density..),binwidth=45,position='identity', alpha = 0.5)+
  theme_classic()+
  scale_fill_manual(values = c("#CE2414","#138EC2"))+
  ylab("Proportion of Values")+
  xlab("Turning Angle (Degrees / s)")+
  xlim(-200,200)

### 2024-03-19 Fish Data Analysis/Results_Text.qmd
Formatting fixes and table

light_no_ab_color = "#d4d7dd"
light_ab_color = "#f5ba9e"
dark_no_ab_color = "#5e94d4"
sim_color = "#777777"

@@

fish_comp_data <- read.csv("Data/Fish_Comp_Values_3D.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"),
                                  Date = paste(Year,Month,Day,Trial, sep = "_"),
                                  Fish_ID = paste(Year,Month,Day,Trial,Fish1, sep = "_")) %>%

                            # ### Used to remove every third point to reduce autocorrelation
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"),
                                  Date = paste(Year,Month,Day,Trial, sep = "_"),
                                  Fish_ID = paste(Year,Month,Day,Trial,Fish1, sep = "_"),
                                  Comp_ID = paste(Year,Month,Day,Trial,Fish1,Fish2, sep = "_")) %>%
                           group_by(Comp_ID) %>%
                           filter(row_number() %% 3 == 1) %>%
                           ungroup()

fish_comp_data <- na.omit(fish_comp_data)

close_min_dist <- 0
close_max_dist <- 3

far_min_dist <- 3
far_max_dist <- 100

fish_data_nn_normal_close <- fish_comp_data %>% filter(Distance >= close_min_dist & Distance <= close_max_dist) %>%
                                        group_by(Flow_Ablation_Darkness) %>%
                                        mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS - mean(Fish1_Yaw_Heading_PS)) %>%
                                        ungroup()

fish_data_nn_normal_far <- fish_comp_data %>% filter(Distance >= far_min_dist & Distance <= far_max_dist) %>%
                                        group_by(Flow_Ablation_Darkness) %>%
                                        mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS - mean(Fish1_Yaw_Heading_PS)) %>%
                                        ungroup()

@@

heading_PS_model_close <- lmer(Fish1_Yaw_Heading_PS ~ sin(Angle) + sin(Angle):Ablation + sin(Angle):Darkness + sin(Angle):Ablation:Darkness + (1|Date), 
                   data = fish_data_nn_normal_close %>% filter(Flow == "Still Water")) 
aov_heading_PS_model_close <- Anova(heading_PS_model_close)

plot(heading_PS_model_close)
qqnorm(resid(heading_PS_model_close))
qqline(resid(heading_PS_model_close))

test(emmeans(heading_PS_model_close, pairwise ~Ablation*Darkness | Angle, at = list(Angle = c(pi/2))))

heading_PS_model_far <- lmer(Fish1_Yaw_Heading_PS ~ sin(Angle) + sin(Angle):Ablation + sin(Angle):Darkness + sin(Angle):Ablation:Darkness + (1|Date) , 
                   data = fish_data_nn_normal_far %>% filter(Flow == "Still Water")) 
aov_heading_PS_model_far <- Anova(heading_PS_model_far)

plot(heading_PS_model_far)
qqnorm(resid(heading_PS_model_far))
qqline(resid(heading_PS_model_far))

test(emmeans(heading_PS_model_far, pairwise ~Ablation*Darkness | Angle, at = list(Angle = c(pi/2))))

turning_school_stats_table <- tibble(
       Near_or_Far = rep(c(rep("Near",4) , rep("Far",4)), 1),
       Condition = rep(c("Angle","Angle x Ablation","Angle x Darkness","Angle x Ablation x Darkness"), 2),
       DF = c(rep(length(heading_PS_model_close@resp$y)-4,4),
                  rep(length(heading_PS_model_far@resp$y)-4,4)),
       Chisq = round(c(aov_heading_PS_model_close$Chisq[1:4],aov_heading_PS_model_far$Chisq[1:4]),3),
       P_Val = round(c(aov_heading_PS_model_close$`Pr(>Chisq)`[1:4],aov_heading_PS_model_far$`Pr(>Chisq)`[1:4]),3),
       Is_Sig = P_Val < 0.05)

>>>>

light_no_ab_color = "#9C9C9C"   #"#C62E65" #A1BAA4" #"#d4d7dd" 06D6A0
light_ab_color = "#f5ba9e"
dark_no_ab_color = "#5e94d4"
dark_ab_color = "#49A078"   #"#1E4371"
sim_color = "#777777"

@@

fish_comp_data <- read.csv("Data/Fish_Comp_Values_3D.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"),
                                  Date = paste(Year,Month,Day,Trial, sep = "_"),
                                  Fish_ID = paste(Year,Month,Day,Trial,Fish1, sep = "_"),
                                  Comp_ID = paste(Year,Month,Day,Trial,Fish1,Fish2, sep = "_"))

min_x <- floor(min(fish_comp_data$Fish1_X, na.rm = T)) + x_offset
max_x <- ceiling(max(fish_comp_data$Fish1_X, na.rm = T)) - x_offset
min_y <- floor(min(fish_comp_data$Fish1_Y, na.rm = T)) + y_offset
max_y <- ceiling(max(fish_comp_data$Fish1_Y, na.rm = T)) - y_offset

print(paste(min_x,max_x,min_y,max_y))

min_dist <- 0
middle_dist <- 2.5
max_dist <- 100

#Limit away from the edge
fish_comp_data <- fish_comp_data %>% filter(Fish1_X >= min_x & Fish1_X <= max_x) %>%
                                     filter(Fish1_Y >= min_y & Fish1_Y <= max_y)

#close
fish_comp_data_ring_angles <- fish_comp_data %>% filter(Distance >= min_dist & Distance <= max_dist) %>%
                                                 mutate(Dist_Cat = ifelse(Distance <= middle_dist, "Close", "Far"))

fish_data_nn_normal <- fish_comp_data_ring_angles %>% group_by(Flow_Ablation_Darkness) %>%
                                        mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS - mean(Fish1_Yaw_Heading_PS, na.rm = T)) %>%
                                        ungroup() %>% 
                                        filter(Flow == "Still Water")

nn_heading_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Angle,pi/6)) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness,Dist_Cat) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup()

@@

perp_angle_df <- fish_data_nn_normal %>% filter(Angle >= deg2rad(75) & Angle <= deg2rad(105))

perp_angle_df_neg <- fish_data_nn_normal %>% filter(Angle >= deg2rad(-105) & Angle <= deg2rad(-75)) %>% 
                                                   mutate(Fish1_Yaw_Heading_PS = -1*Fish1_Yaw_Heading_PS)

perp_angle_df <- rbind(perp_angle_df,perp_angle_df_neg)

perp_angle_df <- perp_angle_df %>% group_by(Flow,Ablation,Darkness,Flow_Ablation_Darkness,Dist_Cat) %>%
                                                mutate(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                       sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                       n = n()) %>%
                                                ungroup()

perp_angle_lmer_close <- lmer(Fish1_Yaw_Heading_PS ~ Ablation*Darkness + (1|Date), data = perp_angle_df %>% filter(Dist_Cat == "Close"))
perp_angle_aov_me_close <- Anova(perp_angle_lmer_close)

plot(perp_angle_lmer_close)
qqnorm(resid(perp_angle_lmer_close))
qqline(resid(perp_angle_lmer_close))

perp_angle_lmer_far <- lmer(Fish1_Yaw_Heading_PS ~ Ablation*Darkness + (1|Date), data = perp_angle_df %>% filter(Dist_Cat == "Far"))
perp_angle_aov_me_far <- Anova(perp_angle_lmer_far)

plot(perp_angle_lmer_far)
qqnorm(resid(perp_angle_lmer_far))
qqline(resid(perp_angle_lmer_far))

turning_school_stats_table <- tibble(
       Near_or_Far = rep(c(rep("Near",3) , rep("Far",3)), 1),
       Condition = rep(c("Ablation","Darkness","Ablation x Darkness"), 2),
       DF = c(rep(length(perp_angle_lmer_close@resp$y)-3,3),
                  rep(length(perp_angle_lmer_far@resp$y)-3,3)),
       Chisq = round(c(perp_angle_aov_me_close$Chisq[1:3],perp_angle_aov_me_far$Chisq[1:3]),3),
       P_Val = round(c(perp_angle_aov_me_close$`Pr(>Chisq)`[1:3],perp_angle_aov_me_far$`Pr(>Chisq)`[1:3]),3),
       Is_Sig = P_Val < 0.05)
